### 2022-04-02 manuscript/thesis.tex
revisions to sections 3 and 4

\begin{table}[h!]
  \centering
  \begin{tabular}{l*{6}{c}r}
  Distribution          & MSD 95\%  \\
  \hline
  Uniform               & 8.809460e-05   \\
  Truncated Normal      & 1.126683e-07  \\
  Truncated Lognormal   & 3.877829e-06  \\
  Truncated Gamma       & 1.500152e-06  \\
  \end{tabular}
  \begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
  \caption[Bin distribution cutoff values]{If a binned 
  one component mixture distribution is fit to a bin forecast
  and the MSD is smaller than
  the corresponding cutoff value listed here, we consider that bin 
  distribution to have been discretized from the fit mixture distribution.}
  \label{table:bincutoffs}
  \end{minipage}
  \end{center}
\end{table}

Of the 869,638 predictions from the CDC Retrospective Forecast project, we fit
each of uniform, TN, TL, and TG distributions to 11,715 of the individual 
predictions. From each of the 27 teams represented, we randomly selected 1,944
forecasts to fit. In many cases the \texttt{optim} function used to fit the 
distributions failed to converge, so of all forecasts we were only able to 
successfully fit distributions to 11,715 of them.
We calculated the MSD, and for each forecast the fit
distribution with the lowest of the four MSDs was considered the best fit. If 
the MSD from the best fit fell
below the corresponding cutoff value listed in \ref{table:bincutoffs} we 
concluded that 
the forecast was discretized from the fit distribution.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/flu_fit_17_10_3.png}}
\caption[Parametric distribution fits to binned distribution]{This figure shows
the plots of fits from each a uniform, truncated normal, truncated lognormal
and truncated gamma distributions to the same binned probability distribution.
In this case, the fit to the normal distribution produced an MSD below the 
normal cutoff value, so we conclude that the binned distribution may have been
discretized froma truncated normal. Proportions of distributions with MSDs 
below the cutoff values are also included.}
\label{fig:binparamfits}
\end{figure}

@@

As of April 1, 2022 there were 89.8 million forecasts  
from 117 different modeling teams submitted to the COVID-19 Forecast Hub (cite??). 
These 
forecasts covered all combinations of 3,202 municipalities (mostly counties) 
in the United States with 441 targets. The first of these forecasts was 
submitted in March of 2020 shortly after the initial outbreak of the COVID-19 
virus in 
the US and forecasts have been received weekly since. 
The forecasts are all quantile forecasts made up of three or 11 predictive 
intervals -depending on the specific time frame for the forecast-
and a median. In the forecasts include seven or 23 quantiles.

@@

A decision was randomly made on creating a quantile distribution with seven 
quantiles or 23
quantiles with probabilities 1/3 and 2/3 respectively. This was done because 
in the COVID-19 Forecast Hub certain target/time frame combinations require 
seven quantiles and others 
require 23. After that decision was made, a random value $u$ was drawn from
a $\mbox{Unif(2,000,\; 25,000)}$ distribution and another value $v$ was drawn
from a $Uniform(3, 200)$ distribution. These values were taken as the mean and
standard deviation and for each of the distribution families considered. The
proper transformations computed to find model parameters corresponding to the 
distribution. Quantile values were calculated for then each quantile. When using 
a t distribution, a value for degrees of freedom was drawn from a 
$\mbox{Unif(2,35)}$ distribution. A CDF
was fit by minimizing \eqref{eq:minthet} over the parameter
vector $\theta$ and the MSE value from \eqref{eq:mse} was 
calculated. The MSE value for which 95\% of the 1,000 simulated distributions
fell below was considered the cutoff for that family and is seen in table 
\ref{table:quantcutoffs}. It is immediately noticeable that the cutoff value 
for the location-scale t family is several orders of magnitude higher than for 
the other distribution families. We believe this is related to the fact that 
that is a three parameter family where as the other four families are two 
parameter families. Thus influencing the optimization methods used enough to 
make such a difference in the cutoff value.

\begin{table}[h!]
  \centering
  \begin{tabular}{l*{6}{c}r}
  Distribution          & Mean Square Error 95\%  \\
  \hline
  Uniform               & 7.220667e-07   \\
  Normal                & 3.097645e-05  \\
  Lognormal             & 1.166775e-07  \\
  Gamma                 & 2.953370e-05  \\
  Location-scale T      & 0.4894471 \\
  \end{tabular}
  \begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
  \caption[Quantile cutoff values]{If a set of quantiles
  is fit to a parametric distribution CDF and the MSE is smaller than
  the corresponding cutoff value listed here, we consider the quantiles to 
  have been calculated parametric distribution.}
  \label{table:quantcutoffs}
  \end{minipage}
  \end{center}
\end{table}

With these cutoff values selected, models were then fit to quantile forecasts
from the COVID-19 Forecasts on zoltardata.com. This was a computationally more
difficult problem than fitting distributions to the binned forecasts of the 
influenza project, so the number of predictions fit is much smaller. The 
predictions fit were selected from each of the 115 models with model weeks, 
targets and units selected randomly. In total 2,504 predictions were fit to each
of the five distributions, the MSE was calculated for each and the fit with the
lowest MSE was selected as the closest fit. If the MSE fell below the cutoff
specified above, we considered that the quantile prediction was approximately
from the same distribution as the fit distribution. Of the 2,504 fits, 99 of 
them produced an MSE value below the corresponding cutoff. Results for fits by
distribution are seen in table \ref{table:cresults}.
Figures \ref{fig:qqfits} and \ref{fig:cdffits} show the QQ plots and CDF plots
for different distribution fits to the same set of quantiles.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/qq_11_12_5.png}}
\caption[QQ plot for quantile fit]{This figure shows qq plots for a set of
submitted quantiles against the quantiles the continuous distribution to which
it was fit. In no case here was the MSE below the cutoff value of the 
corresponding distribution.}
\label{fig:qqfits}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/cdf_11_12_5.png}}
\caption[CDF plot for quantile fit]{This figure shows the CDF plots for 
distributions fit to a set of quantiles. The grey line is a line of connected 
dots from the submitted quantiles and values. 
In no case here was the MSE below the cutoff value of the 
corresponding distribution.}
\label{fig:cdffits}
\end{figure}

The results in the previous two studies suggest that in some cases, forecasters
are using methods which produce simple parametric distribution forecasts. Many
likely are not. To improve upon these studies, one could fit the binned 
influenza forecasts or the COVID-19 quantile forecasts to discrete mixture 
distributions with multiple components. Forecasts that are very close to 
a mixture distribution may suggest that the forecasters already have models that
make transition to a discrete mixture distribution forecast format easy or
straightforward.

@@

In this paper we have reviewed four representation types commonly used in
probabilistic 
forecasting including proper scoring, data storage and ensemble model 
construction for each type. We presented the discrete mixture distribution
representation and we argue that its use in collaborative probabilistic 
forecasting is preferable to the other representations.
In terms of model flexibility, storage and 
ensemble building it is
comparable to discretized bins and interval forecasts but also provides a 
forecast with a infinite nominal resolution. Thus we advocate its use in
future forecasting projects like those done by the CDC and COVID-19 Forecast 
Hub.

For a number of reasons, the adoption of finite mixture distributions as the
submission format may face hurdles. A collaborative forecast center, along with 
forecasters, using a 
different representation may simply not want to break from tradition. The 
implementation of new scoring and ensemble construction methods may be a 
barrier. One aspect of ensemble construction which received little attention in 
this paper is the selection of weights for components of an ensemble where each
of the components is mixture distribution. Computing requirements could become a
concern in such a problem and further research on this may provide
ideas of best methods for weight selection.

Another area of recommended research is the use of joint distributions for 
forecasting. We have only considered here probabilistic forecasting of one 
event at a time. Number of new infections in one week at one specific location
for example. This forecast is presented as a marginal distribution for that 
specific target, time and location. A joint distribution for forecasting 
multiple targets, times or locations may sometime be desirable and may require
further consideration on how joint mixture distributions could be used as a 
format in collaboration.

>>>>

\begin{table}[h!]
  \centering
  \begin{tabular}{l*{6}{c}r}
  Distribution          & MSD 95\% Cutoff \\
  \hline
  Uniform               & 8.809460e-05   \\
  Truncated Normal      & 1.126683e-07  \\
  Truncated Lognormal   & 3.877829e-06  \\
  Truncated Gamma       & 1.500152e-06  \\
  \end{tabular}
  \begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
  \caption[Bin distribution cutoff values]{If a binned 
  one component mixture distribution is fit to a bin forecast
  and the MSD is smaller than
  the corresponding cutoff value listed here, we consider that bin 
  distribution to have been discretized from the fit mixture distribution.}
  \label{table:bincutoffs}
  \end{minipage}
  \end{center}
\end{table}

Of the 869,638 predictions from the CDC Retrospective Forecast project, we fit
each of uniform, TN, TL, and TG distributions to 11,715 of the individual 
predictions. From each of the 27 teams represented, there is a list of forecast
submissions from which we randomly selected 12. From each submission, we 
we randomly selected six forecasts.
In many cases the \texttt{optim} function used to fit the 
distributions failed to converge, so of all forecasts we were only able to 
successfully fit distributions to 11,715 of them.
We calculated the MSD, and for each forecast the fit
distribution with the lowest of the four MSDs was considered the best fit. If 
the MSD from the best fit fell
below the corresponding cutoff value listed in \ref{table:bincutoffs} we 
concluded that 
the forecast was discretized from the fit distribution.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/flu_fit_17_10_3.png}}
\caption[Parametric distribution fits to binned distribution]{This figure shows
the plots of fits from each a uniform, TN, TL,
and TG distributions to the same bin distribution. The title of this forecast
given by the submission team is Bayesian Model Averaging. This forecast was
submitted to the CDC on May 15, 2017 and is a three week ahead forecast of
percent of hospital patients due to an influenza like illness for a specific
region of the United States.
In this case, the fit to the normal distribution produced an MSD below the 
TN cutoff value, so we conclude that the bin forecast may have been
discretized from a TN. Proportions of distributions with MSDs 
below the cutoff values are also included.}
\label{fig:binparamfits}
\end{figure}

@@

As of April 1, 2022 there were 89.8 million forecasts  
from 117 different modeling teams submitted to the COVID-19 Forecast Hub (cite??). 
These 
forecasts covered all combinations of 3,202 municipalities (mostly counties) 
in the United States with 441 targets. The first of these forecasts was 
submitted in March of 2020 shortly after the initial outbreak of the COVID-19 
virus in 
the US and forecasts have been received weekly since. 
The forecasts are all quantile forecasts made up of three or 11 predictive 
intervals -depending on the specific unit for the forecast-
and a median. In the forecasts include seven or 23 quantiles.

@@

A decision was randomly made on creating a quantile distribution with seven 
quantiles or 23
quantiles with probabilities 1/3 and 2/3 respectively. This was done because 
in the COVID-19 Forecast Hub certain target/unit combinations require 
seven quantiles and others 
require 23. After that decision was made, a random value $u$ was drawn from
a $\mbox{Unif(2,000,\; 25,000)}$ distribution and another value $v$ was drawn
from a $\mbox{Unif(3, 200)}$ distribution. These values were taken as the mean 
and
standard deviation and for each of the distribution families considered. The
proper transformations computed to find model parameters corresponding to the 
distribution. Quantile values were calculated for then each quantile. When using 
a t distribution, a value for degrees of freedom was drawn from a 
$\mbox{Unif(2,35)}$ distribution. A CDF
was fit by minimizing \eqref{eq:minthet} over the parameter
vector $\theta$ and the MSE value from \eqref{eq:mse} was 
calculated. The MSE value for which 95\% of the 1,000 simulated distributions
fell below was considered the cutoff for that family and is seen in table 
\ref{table:quantcutoffs}. It is immediately noticeable that the cutoff value 
for the location-scale t family is several orders of magnitude higher than for 
the other distribution families. We believe this is related to the fact that 
that is a three parameter family where as the other four families are two 
parameter families. Thus influencing the optimization methods used enough to 
make such a difference in the cutoff value.

\begin{table}[h!]
  \centering
  \begin{tabular}{l*{6}{c}r}
  Distribution          & MSE 95\% Cutoff \\
  \hline
  Uniform               & 7.220667e-07   \\
  Normal                & 3.097645e-05  \\
  Lognormal             & 1.166775e-07  \\
  Gamma                 & 2.953370e-05  \\
  Location-scale T      & 0.4894471 \\
  \end{tabular}
  \begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
  \caption[Quantile cutoff values]{If a one component mixture distribution of a
  certain distribution family
  is fit to a quantile forecast and the MSE is smaller than
  the corresponding cutoff value listed here, we conclude the quantiles to 
  have been calculated from the fit distribution.}
  \label{table:quantcutoffs}
  \end{minipage}
  \end{center}
\end{table}

With these cutoff values selected, distributions were then fit to quantile 
forecasts
from the COVID-19 Forecasts on zoltardata.com.?? 
From 115 modeling teams we randomly selected four forecast submissions. From 
those submissions we randomly selected three units and attempted to assess
forecasts for all targets under than unit.
This was a computationally more
difficult problem than fitting distributions to the bin forecasts from the CDC
flu competition, so the number of forecasts where convergence for fitting was
met was much smaller. In total distributions from all five families of interest
were fit to 2,504 forecasts. 
The MSE was calculated for each and the fit with the
lowest MSE was selected as the closest fit. If the MSE fell below the cutoff
specified above, we concluded that the quantile forecast was approximately
from the same distribution as the fit distribution.

Table \ref{table:cresults} show the results of the analysis by distribution
family.
Of the 2,504 fits, 99 of 
them had an MSE value below the corresponding cutoff listed in Table
\ref{table:quantcutoffs}. Thus we conclude that a proportion of 0.04 of the 
forecasts assessed came from distributions fit to them.
Results for fits by
distribution are seen in Table \ref{table:cresults}.
Figures \ref{fig:qqfits} and \ref{fig:cdffits} show the QQ plots and CDF plots
for different distribution fits to the same set of quantiles.

@@

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/qq_gleam_110121_1wkincdeath_st16.png}}
\begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
\caption[QQ plot for quantile fit]{This figure shows qq plots for a set of
submitted quantiles against the quantiles the continuous distribution to which
it was fit. 
The name of this forecast given by the team who submitted it is Gleam COVID-19.
This forecast was submitted to the C0VID-19 Forecast Hub on November 1, 2021
and is forecasting deaths one in the next week in a certain US state due to the
virus.
In no case here was the MSE below the cutoff value of the 
corresponding distribution.}
\label{fig:qqfits}
\end{minipage}
\end{center}
\end{figure}

\begin{figure}[htbp]
\centerline{\includegraphics[scale=.15]{Images/cdf_gleam_110121_1wkincdeath_st16.png}}
% i=11,j=12,k=5
\begin{center}
\begin{minipage}{10cm}
\captionsetup{font=scriptsize}
\caption[CDF plot for quantile fit]{This figure shows the CDF plots for 
distributions fit to a set of quantiles. The grey line is a line of connected 
dots from the submitted quantiles and values and is the same in all four plots.
This is the same set of forecasts as is \ref{fig:qqfits}.
In no case here was the MSE below the cutoff value of the 
corresponding distribution.}
\label{fig:cdffits}
\end{minipage}
\end{center}
\end{figure}

The results from these two studies suggest that in some cases, forecasters
are using methods which produce may produce one component mixture distribution
forecasts. Many
likely are not. To expand upon these studies, one could attempt to fit the 
bin forecasts or the quantile forecasts to mixture 
distributions with multiple components. Forecasts that are very close to 
a mixture distribution may suggest that the forecasters already have models that
make transition to a discrete mixture distribution forecast format easy or
straightforward.

@@

In this paper we have reviewed four representation types commonly used in
probabilistic 
forecasting including proper scoring, data storage, and ensemble model 
construction for each type. We presented the mixture distribution
representation and we argue that its use in collaborative probabilistic 
forecasting is preferable to the other representations.
In terms of model flexibility, storage, and 
ensemble building it is
comparable to discretized bins and quantile forecasts but also provides a 
forecast with a infinite nominal resolution. Based on a retrospective analysis,
we argue that among the teams participating in the CDC flu
competition and the COVID-19 Forecast Hub, there are already some which produce
forecasts already resembling single component mixture distributions making the 
transition from past and current formats to a mixture distribution format 
straightforward.
We thus advocate for the use of mixture distributions in
future forecasting projects like those done in the CDC flu competition or in the
COVID-19 Forecast Hub.

For a number of reasons, the adoption of finite mixture distributions as the
submission format may be avoided. A collaborative forecast center, along with 
forecasters, using a 
different representation may simply not want to break from tradition. The 
implementation of new scoring and ensemble construction methods may be a 
barrier. Development of tools beyond what was used in Section 
\ref{section:tools} would assist in making a transition to using mixture 
distributions more straightforward.
One aspect of ensemble construction which received little attention in 
this paper is the selection of weights for components of an ensemble where each
of the components is mixture distribution. Computing requirements could become a
concern in such a problem and further research on this may provide
ideas of best methods for weight selection.

Another area of recommended research is the use of joint mixture distributions 
for 
forecasting. We have only considered here probabilistic forecasting of one 
event at a time. Number of new infections in one week at one specific location
for example. This forecast is presented as a marginal distribution for that 
specific target, time, and location. A joint distribution for forecasting 
multiple targets, times or locations may sometime be desirable and may require
further consideration on how joint mixture distributions could be used as a 
format in collaboration.
